---
title: "Rrra"
author: "David Kaiser"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
sapply(list.files("functions/", full.names = T), source)
library(dplyr)
```

## intro
R toolbox for processing of Radium data from RaDeCC.

## Principal workflow
The main function of this toolbox is `process_samples()`. It accepts a list of files that contains the sample but possibly also other measurement files, data frames containing values for measurement blanks, for detector efficiency values and meta data, as well as some additional arguments.

The files for samples, blanks and standards can be located in the same folder/path. 
```{r}
files <- list.files("data/test_case1/", recursive = TRUE, full.names = TRUE, pattern = ".txt$")
```

The meta data is supplied in a simple table. A wrapper for a read function to access excel sheets will follow soon.
```{r}
meta <- data.frame(id = c("St2", "St3", "St5"),
                   sampling.time = as.POSIXct(c("2021-06-05 09:43:00", "2021-06-05 09:52:00", "2021-06-15 09:52:00")),
                   volume = c(200.4, 200.5, 220.5))
meta
```

Also, we can define those arguments used multiple times.
```{r}
detec = c("orange","blue","grey","green")
datef = "%m/%d/%Y"
standard.id = "standard|std"
blank.id = "blank"
```

The function `summarise_blank()` finds, processes and summarises files containing blank measurements. To identify blanks from the file name, it accepts input to the argument `blank.id`, and to properly summarise, it accepts a vector of detector names in `detectors`. The detector names should be part of the file name. A warning will be printed if less than 3 values are used in a summary. 
```{r}
blk <- summarise_blank(files,
                       date.format = datef, 
                       detectors = detec, 
                       blank.id = blank.id, 
                       summarise = TRUE)
head(blk)
```

Similarly, `summarise_efficiency()` finds and works with files containing standard measurements. It additionally takes input for `standard.id`, as well as DPM values for the standards used. 
```{r}
eff <- summarise_efficiency(files,
                            date.format = datef, 
                            detectors = detec,
                            standard.id = standard.id, 
                            blank.id = blank.id,
                            dpm.223.std = 9.94, # Personal Comm Walter Geibert (AWI)
                            dpm.224.std = 12.1, # Personal Comm Walter Geibert (AWI)
                            summarise = TRUE)
head(eff)
```

Finally, `process_samples()` takes the meta data, output from `summarise_blank()` and `summarise_efficiencies()`, as well as the blank and standard identifiers and an estimate for 227Ac. The detector names are extracted from the inputs to `blk` and `eff`. With `verbose = TRUE` comes a stream of information about file processing, printed to the console. Setting `halfway = TRUE` returns a table of all the values for each sample file that are used for the final calculation of results. All these functions also accept the strptime-style `date.format` to enable correctly converting to POSIXct the Start Time and Stop Time in files created by different system time settings.  
```{r}
out <- process_samples(files = files,
                       blk = blk, 
                       eff = eff, 
                       meta = meta,
                       date.format = datef, 
                       standard.id = standard.id, 
                       blank.id = blank.id,
                       estimate.227Ac = 0.05,
                       halfway = FALSE, 
                       verbose = TRUE)
out
```


## under the hood
### file handling
`files` (paths to files) can be different for samples, blanks and standards, but can also be the same, i.e. all relevant files can be in one folder. Internally, `summarise_blank()`, `summarise_efficiency()` and `process_samples()` call `identify_type()`, which inherits `blank.id` and `standard.id` and uses their values to identify the measurement type from the file name. 
```{r}
identify_type(string = "050621_1grey_St10_blank.txt", blank.id = blank.id, standard.id = standard.id)
identify_type(string = "050621_1orange_228Rastandard.txt", blank.id = blank.id, standard.id = standard.id)
identify_type(string = "060621_blue_223STD_blank.txt", blank.id = blank.id, standard.id = standard.id)
identify_type(string = "St1_224standard_20200223", blank.id = blank.id, standard.id = standard.id)
identify_type(string = "St223_224standard_20200223", blank.id = blank.id, standard.id = standard.id)
```

If neither `blank.id` nor `standard.id` value is found in the string, the measurement is assumed to be from a sample. It is thus not necessary to name samples as such, but is also is no disadvantage.
```{r}
identify_type(string = "050621_1grey_St3.txt", blank.id = blank.id, standard.id = standard.id)
identify_type(string = "050621_1grey_St3.txt", blank.id = "something", standard.id = "other")
```

Only supplying one identifier does work but is not recommended. Only supplying `standard.id` results in a warning, because it is then possible to *mis*identify blanks that were measured before standards (i.e. standard blanks) as standards rather than the blanks they are.
```{r}
identify_type(string = "070621_orange_224STD_blank.txt", standard.id = standard.id)
```

Once the `files` have been filtered for the respective measurement type, the top functions call `read_ra()` on each of those files. `read_ra()` returns a list with all the content of the file.
```{r}
Ra <- read_ra(file = "data/test_case1/samples/050621_1grey_St3.txt", detectors = detec, date.format = datef)
Ra$counts <- head(Ra$counts) # for better readability
Ra
```

The sample ID, `Ra$id` is defined as any string including the (last) occurrence of "St" but before the next "\_" or before the ".txt" extension. This is case-insensitive. For example, `050621_2blue_St2.txt` becomes `St2`, `050621_orange_st1cont.txt` becomes `st1cont`, `050621_st6grey_.txt` becomes `st6grey`. A cruise ID or other info can be tacked on as long as it is not separated by "_": `140621_blue_st5-AlkorCruise_2.txt` becomes `st5-AlkorCruise` but `140621_blue_st5_AlkorCruise_2.txt` becomes `st5`. This created a slight nuisance because standards, that might not have station names, will get a wrong ID, e.g. `20220101_223Standard_2.txt` would become `Standard`. In effect this is not important because the ID of standards is not used in the workflow (other than the samples, who are processed in groups defined by ID).


### data processing
In `summarise_blank()` the outputs of each file are summarised grouped by detector and Ra isotope. In `summarise_blank()` some calculations for detector efficiencies are performed before summarising.  


In `process_samples()`, `process_ra()` takes the output of `read_ra()` and returns values that can be derived solely from the data in one measurement result file.
```{r}
pro <- process_ra(Ra)
t( pro ) # transpose for better readability here
```

`mutate_ra()` takes the output of `summarise_efficiency()`, `summarise_blank()` (inherited from the top function) and `process_ra()`, as well as additional metadata to return more derived values. With the argument `merged.output = TRUE` (the default) the new values will be added to the output of `process_ra()` (this is only useful if the functions are run outside the main workflow).

```{r} 
final <- mutate_ra(eff = eff, 
                   blk = blk,
                   pro = pro,
                   filtration_volume_L = 200.5, 
                   merged.output = TRUE)
t( # transpose for better readability here
  final[!(names(final)%in%names(pro))] # show only those columns not in the output of process_ra()
   )
```

The results are then grouped by the samples´ `id`. 
```{r}
# example count measurements of the same sample
c1 <- read_ra("data/test_case1/samples/050621_1grey_St3.txt", detec, datef) %>% process_ra() %>% mutate_ra(eff, blk, ., 200.5)
c2 <- read_ra("data/test_case1/samples/130621_green_St3_2.txt", detec, datef) %>% process_ra() %>% mutate_ra(eff, blk, ., 200.5)
```

`onFiber_228Th` estimates the amount of 228Th on the fiber to determine the ingrowth of decaying 228Th.
```{r}
onFiber228Th <- onFiber_228Th(midpoint.1 = c1$midpoint, 
                              midpoint.2 = c2$midpoint, 
                              dpm.220per100L.1 = c1$dpm.220per100L,
                              dpm.220per100L.2 = c2$dpm.220per100L)
onFiber228Th
```

The results are then calculated from the processed values of the first measurement/count using `results_Ra()`.
```{r}
results <- results_Ra(count1 = c1, 
                      sampling.time = as.POSIXct("2021-06-05 09:52:00"), 
                      onFiber228Th = onFiber228Th, 
                      estimate.227Ac = 0.05)
results
```

The results of all samples are returned in one data frame.

## Error cases
`read_ra()`: Files with missing ends. Incorrect detector name supplied. Wrong date format supplied.
```{r}
Ra <- read_ra("data/AL557/Count1/050621_1orange_St1.txt", detectors = "pink", date.format = "Y%m%d%")
Ra$counts <- head(Ra$counts)
Ra
```

`identify_type()`: Not supplying either identifier stops the function.
```{r error=TRUE}
identify_type(string = "050621_1grey_St3.txt")
```

`summarise_blank()`: No blanks in `files` stops the function
```{r error=TRUE}
summarise_blank("data/test_case1/samples/", detectors = detec, date.format = filef, blank.id = blank.id)
```

`summarise_efficiency()`: no standards in `files` stops the function
```{r error=TRUE}
summarise_efficiency("data/test_case1/samples/", detectors = detec, date.format = filef, blank.id = blank.id, standard.id = standard.id, dpm.223.std = 10, dpm.224.std = 12)
```

`process_ra()`: missing or incomplete summary stops the function, so does missing Start Time
```{r error=TRUE}
process_ra(read_ra("data/AL557/Count1/050621_1orange_St1.txt", detectors = detec, date.format = datef))
```

`mutate_ra()`: the function stops if the `pro` object's detector has no match in the `blk` or `eff` object

*example coming*

`results_ra()`: the function stops if the times are not supplied correctly or in the correct order
```{r error=TRUE}
results_Ra(count1 = c1, sampling.time = as.POSIXct("2023-12-31 12:12:12"), onFiber228Th = onFiber228Th, estimate.227Ac = 0.05)
```

`process_samples()`: multiple warnings, see above. Wrong meta data formats or no samples in `files` stops the function.
```{r error=TRUE}
process_samples(files = list.files("data/test_case1/blanks/", pattern = ".txt$", full.names = TRUE), 
                blk, eff, meta, datef, blank.id, standard.id, estimate.227Ac = 0.05)
```


# next
- solve this problem:
  - `identify_type(string = "St223_224standard_20200223", blank.id = "blank", standard.id = "standard|std")`
- data output
  - which meta data should be part of the output of process_samples()?
  - should results_Ra() return the decay factors? (currently it doesn't)
  - should summarise_blank() return values for CPMTot? (currently it does but is not used downstream)
- code clean-up 
  - change function and argument names to something nicer
  - get rid of unnecessary ()
  - remove objects not needed downstream
  - unify the use of . and _ in function and argument names
  - remove comments that are notes for the process (e.g. comparisons with excel)
- error handling

# probably dumb questions
- how come sampling time only comes in at results_ra()?
- why use half life of 224Ra when correcting for 228Th decay ingrowth? e.g. in onFiber_228Th()
- in excel sheet blanks cols P and L seem to link to the wrong cols, no downstream effect though.
- it seems that the standards for efficiency are not blank corrected, but there are standard blanks, is this a problem?